A Study of Immersion Factors Using an Augmented Reality Simulation

### 2009-06-24 Paper/paper.tex
created figures directory

\begin{abstract}
It is extremely challenging to run controlled studies comparing multiple augmented reality (AR) systems.  We use an “AR simulation” approach, in which a virtual reality (VR) system is used to simulate multiple AR systems. In order to validate this approach, we carefully replicated a well-known study by Ellis et al. using our simulator, obtaining comparable results.

In a second study using the AR simulator, we studied a tracking task where the AR interface provides x-ray vision to help track a moving target.  We varied the field of view of the AR display, as well as the reliability of the head tracker.  In low reliability conditions, we simulate tracker failure by disabling the augmented view of the scene for brief periods.  Our study gives insight into the effect of head tracker reliability on user performance in a tracking task, as well as the relationship between reliability and field of view in an AR system.

@@

Augmented Reality systems may be used to simulate x-ray vision \cite{1383060}.  The ability to see through solid objects has applications in various fields ranging from engineering and architecture to military or search-and-rescue operations.  Unfortunately, implementing x-ray vision is not a trivial task.  AR systems have cost issules, including those pertaining to level-of-immersion.  For example, what is the ideal augmented field of view (FOV) for a given scenario?  Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors, where tracking is temporarily lost.

@@

motivation??

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=3.5in]{annotated_roomb.png}
	\caption{Overhead view of virtual room (approx. measure).  Participant is stationary in the center of the room, with initial orientation along ``user view'' arrow.  Top-hat's initial position is outside the front doorway, directly in the participant's view.}
\end{figure}

Participants are presented a view of the scene through a head-mounted display helmet with orientation tracking.  The tracking overlays are  augmented on top of the image to form the simulated augmented reality view.

@@

In our experiment we vary two parameters: the field of view of the AR interface and the length of tracking failure periods.  Each trial lasted sixty seconds, and had seven tracking failures.  The total vertical field of view of our HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.  We varied the length of tracking failures between two seconds, one second and zero seconds (no failure).  We experimented with longer values but deemed them too long to be realistic through expert analysis.  This gave us a total of 9 conditions, and for each participant we tested each condition 3 times giving a total of 27 trials per participant.

During each trial there are 20 people in addtition to the man wearing the tophat waking around outside of the room.  The walking path for each person was given by a randomly generated set of points with the constraint that they must stay within the rectangular area of size X by X surrounding the room.  At each point in the path the person would randomly change his walking speed to a new value which was within a range we deemed accecptable through expert analysis.  The man wearing the tophat had the additional constraints as follows.  He must start at the same location for each trial which made him visible by the participant through the door shaped opening in the building.  In order to make the paths by of similar dificulty we decided, through expert analysis, that through the trial he must have walked around the entire building at least once, and he must be visible through the windows for at least $x$ seconds.  We made sure these constraints were met by randomly generating paths until they were satisfied.  We generated 27 sets of patahs this way, one for each trail.  We used a balanced latin square to order these paths and the conditions to discourage the effects of learning.

The experiment was written in python using the Vizard environemnt.  Tracking was done using IntersenseTracker
The HMD we used was a Proview HMD with the following attirbutes. ....  Fill this in.
The computer we used was a Dell with the following specs...........  Fill this in

@@

>>>>

\begin{abstract}
This is a study using a virtual reality (VR) simulation system to examine the effects of various problems associated with augmented reality (AR). In a second study using the AR simulator, we studied a tracking task where the AR interface provides x-ray vision to help track a moving target.  We varied the field of view of the AR display, as well as the reliability of the head tracker.  In low reliability conditions, we simulate tracker failure by disabling the augmented view of the scene for brief periods.  Our study gives insight into the effect of head tracker reliability on user performance in a tracking task, as well as the relationship between reliability and field of view in an AR system.

@@

Augmented Reality systems may be used to simulate x-ray vision \cite{1383060}.  The ability to see through solid objects has applications in various fields ranging from engineering and architecture to military or search-and-rescue operations.  Unfortunately, implementing x-ray vision is not a trivial task.  AR systems have cost issules, including those pertaining to level-of-immersion.  For example, what is the ideal augmented field of view (FOV) for a given scenario?  Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors \cite{4079263}, where tracking is temporarily lost.

@@

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=3.5in]{figures/augmentedroom.png}
	\caption{Overhead view of virtual room (approx. measure).  Participant is stationary in the center of the room, with initial orientation along ``user view'' arrow.  Top-hat's initial position is outside the front doorway, directly in the participant's view.}
\end{figure}

Participants are presented a view of the scene through a head-mounted display helmet with orientation tracking.  The tracking overlays are augmented on top of the original image to form the simulated augmented reality view.

@@

In our experiment we vary two parameters: the field of view of the AR interface and the length of tracking failure periods.  Each trial is 60 seconds, with seven tracking failures.  The total vertical field of view of the HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.  The length of tracking failures varied between two seconds (longest), one second (medium failure) and zero seconds (no failure).  We experimented with longer values but deemed them too long to be realistic through expert analysis.  This gave us a total of 9 conditions, and for each participant we tested each condition 3 times giving a total of 27 trials per participant.

During each trial there are 20 people in addition to a virtual man (Tophat) wearing a tall black tophat walking an unpredictable path outside of the room.  The paths for each virtual person was randomly generated from set of coordinates exterior to the room.  Each path had the constraint that they must stay within the rectangular area of size X by X surrounding the room.  At each path point, the person would randomly change his walking speed to a new value which was within a range we deemed accecptable through expert analysis.  \emph{Tophat} had the additional constraints as follows.  He must start at the same location for each trial which made him visible by the participant through the door shaped opening in the building.  In order to make the paths by of similar dificulty we decided, through expert analysis, that through the trial he must have walked around the entire building at least once, and he must be visible through the windows for at least $x$ seconds.  We made sure these constraints were met by randomly generating paths until they were satisfied.  We generated 27 sets of patahs this way, one for each trail.  We used a balanced latin square to order these paths and the conditions to discourage the effects of learning.

The experiment was written in Python using the Vizard environment.  Tracking was done using InterSense InertiaCube3 orientation tracker which has 180 Hz update rate.  We used a Pro-View 60 head mounted display which displays $640\times480$ video at 60 Hz, and has 36 degrees vertical field of view.  We did not use stereoscopic display.  We ran the experiment on a 2.4 GHz Intel Core 2 machine with 2 Gb of RAM and a NVIDIA GeForce 9800 GX2 graphics card running Windows XP.

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=3.5in]{figures/tophatscreenshot.png}
	\caption{Sample user-view of AR simulation.  \emph{Tophat} is visible (non-occluded) in the doorway.  The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
\end{figure}
